\textbf{Transparency to hybrid open access through publisher-provided metadata: An article-level study of Elsevier}

### 2021-05-09 analysis/manuscript.Rmd
fix grow numbers by invoicing channel

```{r els_waived}
els_waived_df <- hybrid_volume %>% 
  filter(oa_sponsor_type == "ElsevierWaived") 
```

Figure \ref{fig:invoiceoverview} illustrates that over the years, Elsevier has increasingly invoiced hybrid OA APCs directly to authors compared to research funders or academic consortia ("Agreement"). The share of fee-waived articles remained relatively stable, but we found different types of waivers. Around `r filter(els_waived_df, !is.na(oa_sponsor_name)) %>% distinct(doi) %>% nrow() / nrow(els_waived_df)* 100`% of fee waivers were linked to an OA sponsor, which indicates that a third party paid the APC. For instance, `r filter(els_waived_df, oa_sponsor_name == "Académie des Sciences") %>% distinct(doi) %>% nrow()` waivers were linked to the French Académie des Sciences, presumably covering OA publication in its society journals for affiliated authors. The remaining `r filter(els_waived_df, is.na(oa_sponsor_name)) %>% distinct(doi) %>% nrow() / nrow(els_waived_df)*100`% of waived articles did not disclose any OA sponsor. Moreover, Figure \ref{fig:invoiceoverview} compares the invoicing channels based on open content license variants. When Elsevier invoiced authors directly, most OA articles in hybrid journals were published under a non-commercial license (n=`r filter(sponsored_articles, fct_source == "Author", cc_norm == "CC BY-NC-ND") %>% pull(n) %>% sum()`; `r filter(sponsored_articles, fct_source == "Author", cc_norm == "CC BY-NC-ND") %>% pull(n) %>% sum() / filter(sponsored_articles, fct_source == "Author") %>% pull(n) %>% sum()* 100`%), whereas most articles billed as part of agreements were licensed under the more permissive CC BY license (n=`r filter(sponsored_articles, fct_source == "Agreement", cc_norm == "CC BY") %>% pull(n) %>% sum()`; `r filter(sponsored_articles, fct_source == "Agreement", cc_norm == "CC BY") %>% pull(n) %>% sum() / filter(sponsored_articles, fct_source == "Agreement") %>% pull(n) %>% sum() * 100`%).

>>>>

```{r els_waived}
els_waived_df <- hybrid_volume %>% 
  filter(oa_sponsor_type == "ElsevierWaived") 
sponsored_articles_prop <- sponsored_articles %>%
  group_by(issued_year, fct_source) %>%
  summarise(n = sum(n)) %>%
  mutate(perc = n / sum(n) * 100)
```

Figure \ref{fig:invoiceoverview} illustrates that over the years, Elsevier has constantly invoiced hybrid OA APCs directly to authors compared to research funders or academic consortia ("Agreement"). Between 2015 and 2019, the proportion of author-invoicing increased from `r filter(sponsored_articles_prop, issued_year == 2015, fct_source == "Author") %>% pull(perc)`% to `r filter(sponsored_articles_prop, issued_year == 2019, fct_source == "Author") %>% pull(perc)`%, while the share of invoicing through agreements slightly changed from `r filter(sponsored_articles_prop, issued_year == 2015, fct_source == "Agreement") %>% pull(perc)`% to `r filter(sponsored_articles_prop, issued_year == 2019, fct_source == "Agreement") %>% pull(perc)`%. Also, the share of fee-waived articles remained relatively stable, but we found different types of waivers. Around `r filter(els_waived_df, !is.na(oa_sponsor_name)) %>% distinct(doi) %>% nrow() / nrow(els_waived_df)* 100`% of fee waivers were linked to an OA sponsor, which indicates that a third party paid the APC. For instance, `r filter(els_waived_df, oa_sponsor_name == "Académie des Sciences") %>% distinct(doi) %>% nrow()` waivers were linked to the French Académie des Sciences, presumably covering OA publication in its society journals for affiliated authors. The remaining `r filter(els_waived_df, is.na(oa_sponsor_name)) %>% distinct(doi) %>% nrow() / nrow(els_waived_df)*100`% of waived articles did not disclose any OA sponsor. Moreover, Figure \ref{fig:invoiceoverview} compares the invoicing channels based on open content license variants. When Elsevier invoiced authors directly, most OA articles in hybrid journals were published under a non-commercial license (n=`r filter(sponsored_articles, fct_source == "Author", cc_norm == "CC BY-NC-ND") %>% pull(n) %>% sum()`; `r filter(sponsored_articles, fct_source == "Author", cc_norm == "CC BY-NC-ND") %>% pull(n) %>% sum() / filter(sponsored_articles, fct_source == "Author") %>% pull(n) %>% sum()* 100`%), whereas most articles billed as part of agreements were licensed under the more permissive CC BY license (n=`r filter(sponsored_articles, fct_source == "Agreement", cc_norm == "CC BY") %>% pull(n) %>% sum()`; `r filter(sponsored_articles, fct_source == "Agreement", cc_norm == "CC BY") %>% pull(n) %>% sum() / filter(sponsored_articles, fct_source == "Agreement") %>% pull(n) %>% sum() * 100`%).

### 2021-05-09 analysis/manuscript.tex
fix latex compile issue https://github.com/mpark/wg21/issues/54
Making this figure larger (with all panels below each other) to make easier to read
fix grow numbers by invoicing channel

% bibliography
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newenvironment{cslreferences}%
  {\setlength{\parindent}{0pt}%
  \everypar{\setlength{\hangindent}{\cslhangindent}}\ignorespaces}%
  {\par}

\renewcommand{\baselinestretch}{1.5}
\setlength\parindent{24pt}

@@

Figure \ref{fig:invoiceoverview} illustrates that over the years,
Elsevier has increasingly invoiced hybrid OA APCs directly to authors
compared to research funders or academic consortia (``Agreement''). The
share of fee-waived articles remained relatively stable, but we found
different types of waivers. Around 51.7\% of fee waivers were linked to
an OA sponsor, which indicates that a third party paid the APC. For
instance, 853 waivers were linked to the French Académie des Sciences,
presumably covering OA publication in its society journals for
affiliated authors. The remaining 48.3\% of waived articles did not
disclose any OA sponsor. Moreover, Figure \ref{fig:invoiceoverview}
compares the invoicing channels based on open content license variants.
When Elsevier invoiced authors directly, most OA articles in hybrid
journals were published under a non-commercial license (n=32,086;
76.9\%), whereas most articles billed as part of agreements were
licensed under the more permissive CC BY license (n=18,331; 75.6\%).

@@

\caption{\label{tab:subject_profile} Relative subject profile of centrally invoiced OA funding bodies.}
\centering
\resizebox{\linewidth}{!}{
\begin{tabular}[t]{l|rrrrr|rl|rrrrr|rl|rrrrr|rl|rrrrr|rl|rrrrr|rl|rrrrr|rl|rrrrr|r}
\toprule
Country/Subject area & Health science & Life sciences & Physical sciences & SSH & Broad & Total\\
\midrule
United Kingdom & 25\% & 11\% & 21\% & 5\% & 0\% & 62\%\\
The Netherlands & 0\% & 0\% & 0\% & 0\% & 12\% & 12\%\\
European Union & 0\% & 0\% & 0\% & 0\% & 9\% & 9\%\\
United States & 6\% & 0\% & 0\% & 0\% & 1\% & 7\%\\
Norway & 0\% & 0\% & 0\% & 0\% & 3\% & 3\%\\
\addlinespace
Other & 1\% & 0\% & 1\% & 0\% & 5\% & 7\%\\
\midrule
Total & 32\% & 11\% & 22\% & 5\% & 30\% & 100\%\\
\midrule
\bottomrule
\end{tabular}}
\end{table}

>>>>

% bibliography
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newenvironment{CSLReferences}%
  {}%
  {\par}

@@

Figure \ref{fig:invoiceoverview} illustrates that over the years,
Elsevier has constantly invoiced hybrid OA APCs directly to authors
compared to research funders or academic consortia (``Agreement'').
Between 2015 and 2019, the proportion of author-invoicing increased from
57\% to 61.2\%, while the share of invoicing through agreements slightly
changed from 31.6\% to 32.4\%. Also, the share of fee-waived articles
remained relatively stable, but we found different types of waivers.
Around 51.7\% of fee waivers were linked to an OA sponsor, which
indicates that a third party paid the APC. For instance, 853 waivers
were linked to the French Académie des Sciences, presumably covering OA
publication in its society journals for affiliated authors. The
remaining 48.3\% of waived articles did not disclose any OA sponsor.
Moreover, Figure \ref{fig:invoiceoverview} compares the invoicing
channels based on open content license variants. When Elsevier invoiced
authors directly, most OA articles in hybrid journals were published
under a non-commercial license (n=32,086; 76.9\%), whereas most articles
billed as part of agreements were licensed under the more permissive CC
BY license (n=18,331; 75.6\%).

@@

\caption{\label{tab:subject_profile} Relative subject profile of centrally invoiced OA funding bodies.}
\centering
\resizebox{\linewidth}{!}{
\begin{tabular}[t]{lrrrrrr}
\toprule
Country/Subject area & Health science & Life sciences & Physical sciences & SSH & Broad & Total\\
\midrule
United Kingdom & 25\% & 11\% & 21\% & 5\% & 0\% & 62\%\\
The Netherlands & 0\% & 0\% & 0\% & 0\% & 12\% & 12\%\\
European Union & 0\% & 0\% & 0\% & 0\% & 9\% & 9\%\\
United States & 6\% & 0\% & 0\% & 0\% & 1\% & 7\%\\
Norway & 0\% & 0\% & 0\% & 0\% & 3\% & 3\%\\
\addlinespace
Other & 1\% & 0\% & 1\% & 0\% & 5\% & 7\%\\
\midrule
Total & 32\% & 11\% & 22\% & 5\% & 30\% & 100\%\\
\midrule
\bottomrule
\end{tabular}}
\end{table}
